---
title: "CCLE_ParseMutations"
author: "Aurora S Blucher"
date: "2/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

## CCLE Parsing Gene Mutations

Script for parsing selected gene mutations, outputing to file. 
For use with heatmaps and plots - to add gene mutation annotation for key genes like BRCA1/2, TP53, etc.

See CCLE README for mutations; mutations calls are aggregated across sequencing technologies for these cell lines. Using 20Q4 mutation data. Mutation CSV includes mutation call and additional functional calls: TCGA hotspot, COSMIC hotspot, etc.

Notes
02/15/21
-Script created, just running BRCA1/2 for now

02/23/21 
-updates, output BRCA1/2 mutation file  for Alfonso - to  add annotation

```{r}
#read in cell lines - our training set
CCLE_Lines<-read_tsv(here("data", "CCLE_Training", "DepMap_forRepStress_BrEndoOvPanc_TRAINING_020221.tsv"))
View(CCLE_Lines)

#readin mutation data
CCLE_mutation<-read_csv(here("data", "DepmapCCLE_data", "CCLE_mutations.csv"))
View(CCLE_mutation)

#filter to our TRAINING set lines
#and our genes of interest - BRCA1/2 for breast/ovarian, TP53 for ovarian?

BRCA_calls<-CCLE_mutation %>%
  filter(Hugo_Symbol=="BRCA1" | Hugo_Symbol=="BRCA2") %>%
  select(DepMapID = DepMap_ID, Hugo_Symbol, Entrez_Gene_Id, Start_position, End_position, Variant_Type,  Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Protein_Change, isDeleterious, isTCGAhotspot, isCOSMIChotspot)
View(BRCA_calls)

#filter BRCA calls to our list
BRCA_calls_TrainingLines<-BRCA_calls %>%
  filter(DepMapID %in% CCLE_Lines$DepMapID)
View(BRCA_calls_TrainingLines)
#join with cell line information <- for Alfonso training
BRCA_calls_TrainingLinesJoinInfo<-left_join(BRCA_calls_TrainingLines, CCLE_Lines) %>%
  select(DepMapID, CellLine_Name, CellLine_Name_Cleaned, Cancer_Type, Cancer_Subtype, DerivedFrom_Primary_or_Met, Hugo_Symbol, Entrez_Gene_Id, Start_position, End_position, Variant_Type,  Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Protein_Change, isDeleterious, isTCGAhotspot, isCOSMIChotspot)
View(BRCA_calls_TrainingLinesJoinInfo)
#output here for Alfonso 2/23/21
write_tsv(BRCA_calls_TrainingLinesJoinInfo, here("data_intermediate", "CCLEDepMap_TrainingLines_BRCA_Mutations.tsv"))

##check below here -> still compatible?

#split BRCA1 and BRCA2
BRCA1_calls_TrainingLines<-BRCA_calls_TrainingLines %>%
  filter(Hugo_Symbol=="BRCA1")
BRCA2_calls_TrainingLines<-BRCA_calls_TrainingLines %>%
  filter(Hugo_Symbol=="BRCA2")

#okay now take our cell lines and add BRCA annotation as "BRCA1" or "BRCA2"
#output file, 
#feed back into the file for consensus heatmap
CCLE_Lines_annotateBRCA<-CCLE_Lines %>%
  mutate(BRCA_Status_1 = ifelse(DepMapID %in% BRCA1_calls_TrainingLines$DepMapID, "BRCA1", "No"), 
         BRCA_Status_12 = ifelse(DepMapID %in% BRCA2_calls_TrainingLines$DepMapID, "BRCA2", BRCA_Status_1)) %>%
  select(-BRCA_Status_1) #clean-up
View(CCLE_Lines_annotateBRCA)

#output file for consensus heatmap
write_tsv(CCLE_Lines_annotateBRCA, here("data_intermediate", "CCLEDepMap_TrainingLines_BRCA_Annotation.tsv"))

```
