---
title: "DepMapCCLE_ConsensusHeatmap"
author: "Aurora S Blucher"
date: "2/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ComplexHeatmap)
library(circlize) #color scale for heatmaps
```

## R Markdown

Read in the correlation data frames we ran for: McGrail RSR, Peng HRD, McGrail PARPness
Create consensus heatemaps - by disease; across all

Notes
Sun 02/14/21 
-new script just for heatmaps; uses input from Run_RNSSig scripts; cleaned up heatmaps
Mon 02/15/21 
-cleaned up consensus heatmap; z-score, added bar graphs
-parsed subtypes, joined back with our training set (use updated file here*) so we can add to our heatmaps
Tues 02/16/21
-add in BRCA1/2 annotation to heatmap
Fri 02/19/21
-finish BRCA1/2 for ov/panc/end

Sun 04/04/21
-add in a section for RSR Z-score bar plot, where Z-score is across all 4 diseases*
 b/c we need an idea of relative strength here
-fix z-score legends plots - use scale_fill_continuous**

```{r}
#read in correlations for each signature; from other scripts
#add indicator for score type; then go long
RSRScores<-read_tsv(here("data_intermediate", "McGrail_RSR", "CCLEDepMap_TrainingLines_RSR.tsv")) %>%
  mutate(Score = "McGrail_RSR") 
HRDScores<-read_tsv(here("data_intermediate", "Peng_HRD", "CCLEDepMap_TrainingLines_PengHRD.tsv")) %>%
  mutate(Score = "Peng_HRD") 
PARPnessScores<-read_tsv(here("data_intermediate", "McGrail_PARPness", "CCLEDepMap_TrainingLines_McGrailPARPness.tsv")) %>%
  mutate(Score = "McGrail_PARPness")
View(RSRScores)

#bind rows above with Score column then go wide
##add z-score transformation for heatmap, need to scale
#saves us from multiple join and select statements
BindScores<-bind_rows(RSRScores, HRDScores, PARPnessScores) %>%
  group_by(Score) %>%
  mutate(Mean = mean(correlation), SD = sd(correlation), ZScore= (correlation- Mean)/SD) %>%
  dplyr::select(-Mean, -SD, -correlation) %>%
  pivot_wider(names_from = Score, values_from = ZScore) #checked, looks good
View(BindScores)

#read in cell lines - our training set
CCLE_Lines<-read_tsv(here("data", "CCLE_Training", "DepMap_forRepStress_BrEndoOvPanc_TRAINING_020221.tsv"))
#read in our updated file - need the subtype information; added 02/15/20
#left join so we pull in subtypes for our training set ONLY
CCLE_Lines_AddSubtype<-read_tsv(here("data", "DepMap_forRepStress_BrEndoOvPancr_AddSubtypes_021521.txt")) %>%
  dplyr::select(DepMapID, Cancer_Subtype_Specific, Cancer_Molecular_Subtype)
CCLE_LinesJoin_Subtype<-left_join(CCLE_Lines, CCLE_Lines_AddSubtype)

#read in our BRCA1/2 annotation file
CCLE_Lines_BRCA<-read_tsv(here("data_intermediate", "CCLEDepMap_TrainingLines_BRCA_Annotation.tsv")) %>%
  dplyr::select(DepMapID, BRCA_Status_12)

CCLE_LinesJoin<-left_join(CCLE_LinesJoin_Subtype, CCLE_Lines_BRCA)
View(CCLE_LinesJoin)



##BREAST CANCER ONLY
#join with cell line info; updated 02/15 to use
BindScores_CellLineInfo_Br<-left_join(BindScores,CCLE_LinesJoin ) %>% 
  filter(Cancer_Type=="Breast Cancer")
##View(BindScores_CellLineInfo_Br)
#03/14/21- output
write_tsv(BindScores_CellLineInfo_Br, here("data_intermediate", "CellLines_JoinScores", "TrainingLines_JoinInfo_Scores_Breast_BRCA.tsv"))

#set up for heatmap
removeCellLines<-BindScores_CellLineInfo_Br %>%
  dplyr::select(-DepMapID, -CellLine_Name, -CellLine_Name_Cleaned, -Cancer_Type, -Cancer_Subtype,       
        -DerivedFrom_Primary_or_Met, -Cancer_Subtype_Specific, -Cancer_Molecular_Subtype, -BRCA_Status_12)
#View(removeCellLines)
matrix_forHeatMap<-as.matrix(removeCellLines)
rownames(matrix_forHeatMap)<-BindScores_CellLineInfo_Br$CellLine_Name_Cleaned

#row annotation 
#disease type
#primary/metastatic
row_ha = rowAnnotation(Disease = BindScores_CellLineInfo_Br$Cancer_Type, 
                       Derived_from = BindScores_CellLineInfo_Br$DerivedFrom_Primary_or_Met,
                       Subtype = BindScores_CellLineInfo_Br$Cancer_Molecular_Subtype,
                       BRCA_Status = BindScores_CellLineInfo_Br$BRCA_Status_12,
                       col=list(Disease=c("Breast Cancer"="olivedrab3", "Ovarian Cancer"="tomato1", 
                                          "Endometrial/Uterine Cancer" = "darkorchid1", 
                                          "Pancreatic Cancer" = "turquoise"),
                                Derived_from= c("Primary" = "paleturquoise", "Metastasis" = "turquoise4"), 
                                Subtype = c("basal" = "royalblue2", "basal_A" = "royalblue2", "basal_B" = "royalblue2", 
                                            "HER2_amp" ="springgreen4", "luminal"="goldenrod2"),
                                BRCA_Status= c("BRCA1"="orangered2", "BRCA2"="salmon1","No"="gray75" ))
                       )
```

```{r}
#heatmap 
png(filename = here("output", "Consensus_RSR_HRD_PARPness_Breast_Zscore.png"), width = 600, height = 1400)
#works but branches small
Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Z-Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        #column_dend_height = unit(6, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
dev.off()
```

```{r}
#add bar chart - for RSR signature
#idea from McGrail RSR I/O 2021 preprint
#View(BindScores_CellLineInfo_Br)
#02/16/21 - add some modifications to we can add BRCA1/2 status to our barplots
BindScores_CellLineInfo_Br_Annotated<-BindScores_CellLineInfo_Br %>%
  mutate(BRCA_Status = ifelse((BRCA_Status_12=="No"), "No", "Yes"))
#View(BindScores_CellLineInfo_Br_Annotated)

#bar plot for RSR scores
plot<- BindScores_CellLineInfo_Br_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plot + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Breast") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
#add a version of bargraph to show the subtype information
cols<-c("basal" = "royalblue2", "basal_A" = "royalblue2", "basal_B" = "royalblue2", "HER2_amp" ="springgreen4", "luminal"="goldenrod2") 
#bar plot for RSR scores
plotSubtype<- BindScores_CellLineInfo_Br_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR)) +
  geom_bar(stat="identity", aes(fill=Cancer_Molecular_Subtype, color=BRCA_Status)) + 
  scale_fill_manual(values = cols, na.value = "grey")  +
  scale_color_manual(values = c("No"=NA, "Yes"="orangered2"))
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Breast") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  

```


```{r}
#Repeat for Ovarian
BindScores_CellLineInfo_Ov<-left_join(BindScores,CCLE_LinesJoin  ) %>% 
  filter(Cancer_Type=="Ovarian Cancer")
##View(BindScores_CellLineInfo_Ov)
#03/14/21- output
write_tsv(BindScores_CellLineInfo_Ov, here("data_intermediate", "CellLines_JoinScores", "TrainingLines_JoinInfo_Scores_Ov_BRCA.tsv"))

#set up for heatmap
removeCellLines<-BindScores_CellLineInfo_Ov %>%
  dplyr::select(-DepMapID, -CellLine_Name, -CellLine_Name_Cleaned, -Cancer_Type, -Cancer_Subtype, 
         -DerivedFrom_Primary_or_Met, -Cancer_Subtype_Specific, -Cancer_Molecular_Subtype,-BRCA_Status_12)

matrix_forHeatMap<-as.matrix(removeCellLines)
rownames(matrix_forHeatMap)<-BindScores_CellLineInfo_Ov$CellLine_Name_Cleaned

#row annotation 
#disease type
#primary/metastatic
row_ha = rowAnnotation(Disease = BindScores_CellLineInfo_Ov$Cancer_Type, 
                       Derived_from = BindScores_CellLineInfo_Ov$DerivedFrom_Primary_or_Met,
                       Subtype = BindScores_CellLineInfo_Ov$Cancer_Subtype_Specific,
                       BRCA_Status = BindScores_CellLineInfo_Ov$BRCA_Status_12,
                       col=list(Disease=c("Breast Cancer"="olivedrab3", "Ovarian Cancer"="tomato1", 
                                          "Endometrial/Utering Cancer" = "darkorchid1", 
                                          "Pancreatic Cancer" = "turquoise"),
                                Derived_from= c("Primary" = "paleturquoise", "Metastasis" = "turquoise4"), 
                                Subtype = c("clear_cell"="thistle2", "endometrioid"="springgreen4", "high_grade_serous"="royalblue2", "low_grade_serous"="steelblue2", "mixed_endometrioid_clear_cell"="plum1", "mixed_serous_clear_cell"="plum1", "mucinous" ="darkgoldenrod2" , "serous"="tan"),
                                BRCA_Status= c("BRCA1"="orangered2", "BRCA2"="salmon1","No"="gray75" ))
                       )
```

```{r}
#heatmap 
png(filename = here("output", "Consensus_RSR_HRD_PARPness_Ovarian_Zscore.png"), width = 600, height = 1400)
#size to show dendograms and labels, make sure resolution is good when you zoom
Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Z-Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        #column_dend_height = unit(6, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
dev.off()
```

```{r}
#02/19/21 - add some modifications to we can add BRCA1/2 status to our barplots
BindScores_CellLineInfo_Ov_Annotated<-BindScores_CellLineInfo_Ov %>%
  mutate(BRCA_Status = ifelse((BRCA_Status_12=="No"), "No", "Yes"))
#View(BindScores_CellLineInfo_Br_Annotated)

plotOv<- BindScores_CellLineInfo_Ov%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato") 
plotOv + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Ovarian") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#add ovarian subtype  
#add a version of bargraph to show the subtype information
cols<-c("clear_cell"="thistle2", "endometrioid"="springgreen4", "high_grade_serous"="royalblue2", "low_grade_serous"="steelblue2", "mixed_endometrioid_clear_cell"="plum1", "mixed_serous_clear_cell"="plum1", "mucinous" ="darkgoldenrod2" , "serous"="tan" )
#bar plot for RSR scores
plotSubtypeOv<- BindScores_CellLineInfo_Ov_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=Cancer_Subtype_Specific, color=BRCA_Status)) + 
  scale_fill_manual(values = cols, na.value = "grey")  +
  scale_color_manual(values = c("No"=NA, "Yes"="orangered2"))
plotSubtypeOv + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Ovarian") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="subtype")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))
 

```


```{r}
#Repeat for Endometrial
BindScores_CellLineInfo_Endo<-left_join(BindScores,CCLE_LinesJoin  ) %>% 
  filter(Cancer_Type=="Endometrial/Uterine Cancer")
#View(BindScores_CellLineInfo_Endo)
#03/14/21- output
#write_tsv(BindScores_CellLineInfo_Endo, here("data_intermediate", "CellLines_JoinScores", "TrainingLines_JoinInfo_Scores_Endo_BRCA.tsv"))


#set up for heatmap
removeCellLines<-BindScores_CellLineInfo_Endo%>%
  dplyr::select(-DepMapID, -CellLine_Name, -CellLine_Name_Cleaned, -Cancer_Type, -Cancer_Subtype, -DerivedFrom_Primary_or_Met,
         -Cancer_Subtype_Specific, -Cancer_Molecular_Subtype, -BRCA_Status_12)

matrix_forHeatMap<-as.matrix(removeCellLines)
rownames(matrix_forHeatMap)<-BindScores_CellLineInfo_Endo$CellLine_Name_Cleaned

#row annotation 
#disease type
#primary/metastatic
row_ha = rowAnnotation(Disease = BindScores_CellLineInfo_Endo$Cancer_Type, 
                       Derived_from = BindScores_CellLineInfo_Endo$DerivedFrom_Primary_or_Met,
                       Subtype= BindScores_CellLineInfo_Endo$Cancer_Subtype,
                       BRCA_Status = BindScores_CellLineInfo_Endo$BRCA_Status_12,
                       col=list(Disease=c("Breast Cancer"="olivedrab3", "Ovarian Cancer"="tomato1", 
                                          "Endometrial/Uterine Cancer" = "darkorchid1", 
                                          "Pancreatic Cancer" = "turquoise"),
                                Derived_from= c("Primary" = "paleturquoise", "Metastasis" = "turquoise4"), 
                                Subtype=c("Choriocarcinoma"="lightsalmon1", "Clear Cell Carcinoma"="orange2",
                                          "Endometrial Adenocarcinoma"="royalblue2" ,
                                          "Endometrial Adenosquamous Carcinoma"="springgreen2",
                                          "Endometrial Squamous Cell Carcinoma"="springgreen4",
                                          "Endometrial Stromal Sarcoma"="maroon1",
                                          "Mullerian Carcinoma"="goldenrod2",
                                          "Uterine Carcinosarcoma"="slateblue1"),
                                BRCA_Status= c("BRCA1"="orangered2", "BRCA2"="salmon1","No"="gray75" ) )
                       )
```

```{r}
#heatmap 
png(filename = here("output", "Consensus_RSR_HRD_PARPness_Endometrial_ZScore.png"), width = 800, height = 1400)
#size to show dendograms and labels, make sure resolution is good when you zoom
Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Z-Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        #column_dend_height = unit(6, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
dev.off()
```

```{r}
#add BRCA status
BindScores_CellLineInfo_Endo_Annotated<-BindScores_CellLineInfo_Endo %>%
  mutate(BRCA_Status = ifelse((BRCA_Status_12=="No"), "No", "Yes"))
#View(BindScores_CellLineInfo_Endo_Annotated)

plotEndo<- BindScores_CellLineInfo_Endo%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotEndo + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Endometrial") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
#add endometrial subtype barplot
#add a version of bargraph to show the subtype information
cols<-c("Choriocarcinoma"="lightsalmon1", "Clear Cell Carcinoma"="orange2",
                                          "Endometrial Adenocarcinoma"="royalblue2" ,
                                          "Endometrial Adenosquamous Carcinoma"="springgreen2",
                                          "Endometrial Squamous Cell Carcinoma"="springgreen4",
                                          "Endometrial Stromal Sarcoma"="maroon1",
                                          "Mullerian Carcinoma"="goldenrod2",
                                          "Uterine Carcinosarcoma"="slateblue1")
#bar plot for RSR scores
plotSubtypeEndo<- BindScores_CellLineInfo_Endo_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=Cancer_Subtype, color=BRCA_Status)) + 
  scale_fill_manual(values = cols, na.value = "grey") +
  scale_color_manual(values = c("No"=NA, "Yes"="orangered2"))
plotSubtypeEndo + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Endometrial") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 8))
 

```

```{r}
#Repeat for Pancreatic
##no subtype breakoutneeded - all are Ductal Adenocarcinoma_exocrine
##except for QGP1(ACH-000347) = Somatostatinoma
BindScores_CellLineInfo_Panc<-left_join(BindScores, CCLE_LinesJoin) %>% 
  filter(Cancer_Type=="Pancreatic Cancer")
#View(BindScores_CellLineInfo_Panc)
#03/14/21- output
#write_tsv(BindScores_CellLineInfo_Panc, here("data_intermediate", "CellLines_JoinScores", "TrainingLines_JoinInfo_Scores_Panc_BRCA.tsv"))


#set up for heatmap
removeCellLines<-BindScores_CellLineInfo_Panc%>%
  dplyr::select(-DepMapID, -CellLine_Name, -CellLine_Name_Cleaned, -Cancer_Type, -Cancer_Subtype,-Cancer_Subtype_Specific, -Cancer_Molecular_Subtype, -DerivedFrom_Primary_or_Met, -BRCA_Status_12)
View(removeCellLines)

matrix_forHeatMap<-as.matrix(removeCellLines)
rownames(matrix_forHeatMap)<-BindScores_CellLineInfo_Panc$CellLine_Name_Cleaned

#row annotation 
#disease type
#primary/metastatic
row_ha = rowAnnotation(Disease = BindScores_CellLineInfo_Panc$Cancer_Type, 
                       Derived_from = BindScores_CellLineInfo_Panc$DerivedFrom_Primary_or_Met,
                       BRCA_Status = BindScores_CellLineInfo_Panc$BRCA_Status_12,
                       col=list(Disease=c("Breast Cancer"="olivedrab3", "Ovarian Cancer"="tomato1", 
                                          "Endometrial/Uterine Cancer" = "darkorchid1", 
                                          "Pancreatic Cancer" = "turquoise"),
                                Derived_from= c("Primary" = "paleturquoise", "Metastasis" = "turquoise4"),
                                BRCA_Status= c("BRCA1"="orangered2", "BRCA2"="salmon1","No"="gray75" ))
                       )
```

```{r}
#add BRCA status
BindScores_CellLineInfo_Panc_Annotated<-BindScores_CellLineInfo_Panc %>%
  mutate(BRCA_Status = ifelse((BRCA_Status_12=="No"), "No", "Yes"))
#View(BindScores_CellLineInfo_Panc_Annotated)

plotPanc<- BindScores_CellLineInfo_Panc_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) +
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotPanc + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Pancreatic") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#add BRCA annotation
plotPanc<- BindScores_CellLineInfo_Panc_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR, color=BRCA_Status)) +
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotPanc + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Pancreatic") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

```{r}
#heatmap 
png(filename = here("output", "Consensus_RSR_HRD_PARPness_Panc_ZScore.png"), width = 800, height = 1400)
#size to show dendograms and labels, make sure resolution is good when you zoom
Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Z-Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        #column_dend_height = unit(6, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
dev.off()
```

```{r}
#HEATMAP OF CONSENSUS ACROSS ALL
#Reviewed 04/04/21
# Our Z-score was calculated across ALL diseases, so this is relative
# Add barplots for within-calculated z-scores below
BindScores_CellLineInfo_All<-left_join(BindScores,CCLE_LinesJoin ) 
#View(BindScores_CellLineInfo_All)

#set up for heatmap
removeCellLines<-BindScores_CellLineInfo_All %>%
  dplyr::select(-DepMapID, -CellLine_Name, -CellLine_Name_Cleaned, -Cancer_Type, -Cancer_Subtype, -DerivedFrom_Primary_or_Met,
         -Cancer_Subtype_Specific, -Cancer_Molecular_Subtype, -BRCA_Status_12)

matrix_forHeatMap<-as.matrix(removeCellLines)
rownames(matrix_forHeatMap)<-BindScores_CellLineInfo_All$CellLine_Name_Cleaned

#row annotation 
#disease type
#primary/metastatic
row_ha = rowAnnotation(Disease = BindScores_CellLineInfo_All$Cancer_Type, 
                       Derived_from = BindScores_CellLineInfo_All$DerivedFrom_Primary_or_Met,
                       BRCA_Status = BindScores_CellLineInfo_All$BRCA_Status_12,
                       col=list(Disease=c("Breast Cancer"="olivedrab3", "Ovarian Cancer"="tomato1", 
                                          "Endometrial/Uterine Cancer" = "darkorchid1", 
                                          "Pancreatic Cancer" = "turquoise"),
                                Derived_from= c("Primary" = "paleturquoise", "Metastasis" = "turquoise4"),
                                BRCA_Status= c("BRCA1"="orangered2", "BRCA2"="salmon1","No"="gray75" ))
                       )
```

```{r}
#heatmap 
png(filename = here("output", "Consensus_RSR_HRD_PARPness_All_Zscore_AddBRCA.png"), width = 1000, height = 2400)
#size to show dendograms and labels, make sure resolution is good when you zoom
Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Z-Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        #column_dend_height = unit(6, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
dev.off()
```

```{r}
plotAll<- BindScores_CellLineInfo_All%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=Cancer_Type))  
  #scale_fill_gradient(low = "dodgerblue2", high = "tomato") 
plotAll + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Br/Ov/Endo/Panc") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))
  
#Show BRCA1/2 status
BindScores_CellLineInfo_All_Annotated<-BindScores_CellLineInfo_All %>%
  mutate(BRCA_Status = ifelse((BRCA_Status_12=="No"), "No", "Yes"))
View(BindScores_CellLineInfo_All_Annotated)

plotAll<- BindScores_CellLineInfo_All_Annotated%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=BRCA_Status))+
  scale_fill_manual(values = c("No"="grey75", "Yes"="orangered2"))
plotAll + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Br/Ov/Endo/Panc") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="BRCA Mutated")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))
 
```
```{r}
#Add section for disease-relative z-scores
#and run the bar plots again
BindScores_forZScoreRelative<-bind_rows(RSRScores, HRDScores, PARPnessScores)
#View(BindScores_forZScoreRelative)
  
#join with the cell line info
BindScores_CellLineInfo_forZScoreRelative<-left_join(BindScores_forZScoreRelative,CCLE_LinesJoin)
BindScores_CellLineInfo_forZScoreRelative_RSR<-BindScores_CellLineInfo_forZScoreRelative %>%
  filter(Score=="McGrail_RSR")
#View(BindScores_CellLineInfo_forZScoreRelative_RSR)

RSR_ZScore_WithinDiseaseGroup<-BindScores_CellLineInfo_forZScoreRelative_RSR %>%
  group_by(Cancer_Type) %>%
  mutate(Mean = mean(correlation), SD = sd(correlation), ZScore= (correlation- Mean)/SD) %>%
  dplyr::select(-Mean, -SD, -correlation)
View(RSR_ZScore_WithinDiseaseGroup)
  
#now add barplots for each disease - relative here
#note how the shape of the distribution changes
#because it is calculated within group
plotBr_Relative<- RSR_ZScore_WithinDiseaseGroup%>%
  filter(Cancer_Type=="Breast Cancer") %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -ZScore), y=ZScore, fill=ZScore),
         color="black") +
  geom_bar(stat="identity")+#, aes(fill=Cancer_Type))  
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotBr_Relative + labs(title = "Distribution of McGrail RSR Score (Z-Scores) Calculated WITHIN Disease Group", subtitle = "CCLE Breast") +
  xlab("") +
  ylab("Z-Score") +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))#+
  #facet_wrap(~Cancer_Type, nrow=2, ncol=2)

plotOv_Relative<- RSR_ZScore_WithinDiseaseGroup%>%
  filter(Cancer_Type=="Ovarian Cancer") %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -ZScore), y=ZScore, fill=ZScore),
         color="black") +
  geom_bar(stat="identity")+#, aes(fill=Cancer_Type))  
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotOv_Relative + labs(title = "Distribution of McGrail RSR Score (Z-Scores) Calculated WITHIN Disease Group", subtitle = "CCLE Ovarian") +
  xlab("") +
  ylab("Z-Score") +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))#+
  #facet_wrap(~Cancer_Type, nrow=2, ncol=2)

plotEnd_Relative<- RSR_ZScore_WithinDiseaseGroup%>%
  filter(Cancer_Type=="Endometrial/Uterine Cancer") %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -ZScore), y=ZScore, fill=ZScore),
         color="black") +
  geom_bar(stat="identity")+#, aes(fill=Cancer_Type))  
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotEnd_Relative + labs(title = "Distribution of McGrail RSR Score (Z-Scores) Calculated WITHIN Disease Group", subtitle = "CCLE Endometrial") +
  xlab("") +
  ylab("Z-Score") +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))#+
  #facet_wrap(~Cancer_Type, nrow=2, ncol=2)

plotPanc_Relative<- RSR_ZScore_WithinDiseaseGroup%>%
  filter(Cancer_Type=="Pancreatic Cancer") %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -ZScore), y=ZScore, fill=ZScore),
         color="black") +
  geom_bar(stat="identity")+#, aes(fill=Cancer_Type))  
  scale_fill_continuous(low = "dodgerblue2", high = "tomato", name="Z-Score") 
plotPanc_Relative + labs(title = "Distribution of McGrail RSR Score (Z-Scores) Calculated WITHIN Disease Group", subtitle = "CCLE Pancreatic") +
  xlab("") +
  ylab("Z-Score") +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=2.5))#+
  #facet_wrap(~Cancer_Type, nrow=2, ncol=2)

```
