---
title: "ConvertHumanToMouse"
author: "Aurora S Blucher"
date: "2/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(BiocManager)# if needed, run #BiocManager::install(version = "3.12")
#BiocManager::install("biomaRt")
library(biomaRt)
#package doc http://www.bioconductor.org/packages/release/bioc/manuals/biomaRt/man/biomaRt.pdf
```

## Converting Human Gene Symbols to Mouse Gene Symbols Using BioMart

Quick script for converting from human gene symbols to mouse gene symbols. Queries BioMart software for Ensembl database and human/mouse datasets. Note that for some human gene IDs in the RSR signature, we pull back multiple mouse gene IDs. For a first pass we are keeping all of these.

Notes
02/17/21
-convert  RSR gene list (human) to mouse so we can run for our MDST models
02/18/21
-add McGrail PARPness, Peng HRD signatures as well

```{r}
#McGrail RSR Genes and Coefficients
RSRSignatureGenes<-read_tsv(here("data","RNA_Signatures", "McGrailRSR_Genes_Coefficients.tsv")) %>%
  arrange(Gene)
#View(RSRSignatureGenes) #711 genes total

#McGrail PARPness Genes and Coefficients
PARPnessSignatureGenes<-read_tsv(here("data","RNA_Signatures", "McGrailPARPness_Genes_Coefficients.tsv")) 
#View(PARPnessSignatureGenes)
#Peng HRD Genes and Coefficients
HRDSignatureGenes<-read_tsv(here("data","RNA_Signatures", "PengHRDGenes_BRCA1_Coefficients.tsv"))
#View(HRDSignatureGenes)


#Query Biomart for Biomart database (Ensembl) and human/mouse datasets
#Load human/ mouse genes
HumanGenes = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
MouseGenes = useMart("ensembl", dataset = "mmusculus_gene_ensembl")

# #test example
# testGenesHuman<-c("ABCB6", "ABCC2", "CEBPD")
# 
# #run getLDS(), values should be a vector
# genesConvert<-getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values=testGenesHuman, mart = HumanGenes, attributesL= c("mgi_symbol"), martL =MouseGenes, uniqueRows = T )
# View(genesConvert) #looks good

#MCGRAIL RSR
#convert the RSR genes
genesConvert<-getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values=RSRSignatureGenes$Gene, mart = HumanGenes, attributesL= c("mgi_symbol"), martL =MouseGenes, uniqueRows = T )
View(genesConvert) #609 genes converted

#join back with the coefficients
McGrailRSR_ConvertedToMouse<-genesConvert %>%
  dplyr::select(Gene = HGNC.symbol, MGI.symbol) #use dplyr::select because there is a conflict with another pkg select() call
McGrailRSR_ConvertedToMouseJoin<-left_join(McGrailRSR_ConvertedToMouse, RSRSignatureGenes)
View(McGrailRSR_ConvertedToMouseJoin)

#note, in some cases the human gene ID  maps to multiple mouse gene IDS
#for now we are keeping all of these

#output to file
write_tsv(McGrailRSR_ConvertedToMouseJoin, 
          here("data_intermediate", "MDST", "McGrailRSR_Genes_Coefficients_ConvertedToMouse.tsv"))

#MCGRAIL PARPness
#convert the RSR genes
genesConvertParpness<-getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values=PARPnessSignatureGenes$Gene, mart = HumanGenes, attributesL= c("mgi_symbol"), martL =MouseGenes, uniqueRows = T )
View(genesConvertParpness) #130 genes converted

#join back with the coefficients
McGrailParpness_ConvertedToMouse<-genesConvertParpness %>%
  dplyr::select(Gene = HGNC.symbol, MGI.symbol) #use dplyr::select because there is a conflict with another pkg select() call
McGrailParpness_ConvertedToMouseJoin<-left_join(McGrailParpness_ConvertedToMouse, PARPnessSignatureGenes)
View(McGrailParpness_ConvertedToMouseJoin)

#output to file
write_tsv(McGrailParpness_ConvertedToMouseJoin, 
          here("data_intermediate", "MDST", "McGrailPARPness_Genes_Coefficients_ConvertedToMouse.tsv"))

#PENG HRD
#convert the RSR genes
genesConvertHRD<-getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values=HRDSignatureGenes$Gene, mart = HumanGenes, attributesL= c("mgi_symbol"), martL =MouseGenes, uniqueRows = T )
View(genesConvertHRD) 

#join back with the coefficients
McGrailParpness_ConvertedToMouse<-genesConvertParpness %>%
  dplyr::select(Gene = HGNC.symbol, MGI.symbol) #use dplyr::select because there is a conflict with another pkg select() call
McGrailParpness_ConvertedToMouseJoin<-left_join(McGrailParpness_ConvertedToMouse, PARPnessSignatureGenes)
View(McGrailParpness_ConvertedToMouseJoin)

#output to file
write_tsv(McGrailParpness_ConvertedToMouseJoin, 
          here("data_intermediate", "MDST", "McGrailPARPness_Genes_Coefficients_ConvertedToMouse.tsv"))


```
