---
title: "RunGSVA"
author: "Aurora S Blucher"
date: "3/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("msigdbr")
library(msigdbr)
library(here)
library(tidyverse)
library(ggpubr)
library(ComplexHeatmap)
```

## Run GSVA

Notes for developing the script for running GSVA on RNASeq data. The Molecular Signatures Database (MSigDB) houses genes sets that were created for use with GSEA (Gene Set Enrichment Analysis). Here we are interested in using the focused set of Cancer Hallmarks (50 gene sets) with GSVA (Gene Set Variation Analysis), and in particular these genee sets of interest for replication stress:
-G2M_CHECKPOINT
-E2F_TARGETS
-MYC_TARGETS
-DNA Repair
-OxPhos ~maybe

TODO: Meet and discuss with Allie

Notes 
Mon 03/01/21
-get msigDB hallmarks sets
-convert MDST expression into eSet
-set up GSVA

Thurs 03/11/21-Sun 03/14/21*
-running GSVA on MDST models; output ES scores; create heatmap for selectd pathways
-overall heatmap - all 50 hallmarks
-rep stress focused heatmap-> for slides




```{r}
#MSIGDB CANCER HALLMARKS
#following documentation here
#https://cran.r-project.org/web/packages/msigdbr/vignettes/msigdbr-intro.html

#see all the msigdbr collctions
#msigdbr_collections()

hallmark_GeneSets<-msigdbr(species = "Mus musculus", category = "H")
class(hallmark_GeneSets)
View(hallmark_GeneSets)

#number of hallmarks -  should be 50
View(unique(hallmark_GeneSets$gs_name)) #50 total

#select hallmarks of interest and get our list
#G2M_CheckPoints, E2F

RepStressSets_OfInterest<-hallmark_GeneSets%>%
  filter(gs_name=="HALLMARK_G2M_CHECKPOINT"|gs_name=="HALLMARK_E2F_TARGETS" | gs_name=="HALLMARK_DNA_REPAIR" | 
         gs_name=="HALLMARK_MYC_TARGETS_V1" | gs_name=="HALLMARK_MYC_TARGETS_V2")
View(RepStressSets_OfInterest) #808 entries, but likely contains overlap

View(unique(RepStressSets_OfInterest$gene_symbol)) #645 unique genes, for human #644 for mouse

#so interestinly the RSR set and the GSVA set are both very large
#how much overlap in these? *compute
#check overlap, unique
#check correlations 
#
#

#syntax for using gene sets in GSVA
#msigdbr_list = split(x = msigdbr_df$gene_symbol, f = msigdbr_df$gs_name)
#gsva(gset.idx.list = msigdbr_list, ...)

#what does this do?
msigdbr_list = split(x = hallmark_GeneSets$gene_symbol, f = hallmark_GeneSets$gs_name)
class(msigdbr_list)#list
View(msigdbr_list) 
#contains 1 list for each hallmark-> gene sets
class(msigdbr_list$HALLMARK_G2M_CHECKPOINT) #character?
View(msigdbr_list$HALLMARK_G2M_CHECKPOINT)#data frame 1 column/ vector of all the genes in this set here
#gsva(gset.idx.list = msigdbr_list, ...)

```

```{r}
###SET UP
#expression set syntax
#following bioconductor documentation here #https://www.bioconductor.org/packages/release/bioc/vignettes/Biobase/inst/doc/ExpressionSetIntroduction.pdf

#we need to load our MDST data into an expression set
library(BiocManager)
library(Biobase)
library(readxl)
#load in our MDST data; RNASeq matrix
MDST_RNASeq_dataFrame<-read_excel(here("data", "MDST_Models", "MDST_RNASeq_cpm_annotated.xlsx"), range = "A6:AA16186") %>%
  dplyr::select(-gene_ID, -...3)
head(MDST_RNASeq_dataFrame)
View(MDST_RNASeq_dataFrame)

class(MDST_RNASeq_dataFrame$Gene_name) #char
class(MDST_RNASeq_dataFrame$Control1_FatPad_F2) #numeric -> wonder what is causing the problem

#convert to matrix
exprs<-as.matrix(MDST_RNASeq_dataFrame[, 2:25]) #NEED TO SKIP the first column of genes
class(exprs)
dim(exprs) #16180 genes X 24 samples (mice)
colnames(exprs)#MDST names
rownames(exprs)<-t(MDST_RNASeq_dataFrame[,1])# not set? #added tues

head(exprs)

#create a minimal ExpressionSet object
minimalSet<-ExpressionSet(assayData=exprs[1:100, 1:5])
minimalSet
View(minimalSet)
#if we wanted to set phenotypic data then we would do so with 
#pData<-dataFramePhenotypicInfo

#exploratory GSVA work
#following pg 8 here
#https://www.bioconductor.org/packages/release/bioc/vignettes/GSVA/inst/doc/GSVA.pdf

library(GSVA)
#feed in our expressionSet
#feed in cancer hallmarks as the msigdbr_list object, list of gene set name -> genes; see if that works
#no annotation arg; "in the case of calling gsva() on a ExpressionSet object, the annotation argument is ignored
test_gsva<-gsva(minimalSet,gset.idx.list=msigdbr_list, method=c("gsva"), verbose=TRUE) #testing 03/01/21
test_gsva
#did not work / ran for test sample set!

#okay, how do we access the data
#unclear
test_gsva
featureNames(test_gsva)#features = hallmarks, 39 ran here
sampleNames(test_gsva) #samples we ran
experimentData(test_gsva)#not informative

#extract th expression matrix -> should this now be 
testMatrixES<-exprs(test_gsva)
View(testMatrixES)

#if you want to subset, it is similar to subsetting the expression matrix contained inside ExpressionSet
subsetMatrixES<-test_gsva[1:5, 1:3]
dim(subsetMatrixES) #5 features, 3 samples
featureNames(subsetMatrixES) #5 hallmarks
sampleNames(subsetMatrixES) #3 samples
seeSubsetMatrix<-exprs(subsetMatrixES) #get just the expression matirx -> which here is enrichment scores
View(seeSubsetMatrix)

```

```{r}
#03/14/21 - MDST run GSVA and create heatmaps

#MDST data frame, remove duplicate row namese
#not sure why we have 2 listings for these genes
MDST_RNASeq_dataFrame_RemoveDups<-MDST_RNASeq_dataFrame %>%
  select(Gene_name) %>%
  group_by(Gene_name) %>%
  tally() %>%
  filter(n==1)
View(MDST_RNASeq_dataFrame_RemoveDups)

#restrict our full data frame to just our genes with only 1 count
MDST_RNASeq_dataFrame_cleaned<-MDST_RNASeq_dataFrame %>%
  filter(Gene_name %in% MDST_RNASeq_dataFrame_RemoveDups$Gene_name)
View(MDST_RNASeq_dataFrame_cleaned)

#run for our full MDST data frame
exprsFull<-as.matrix(MDST_RNASeq_dataFrame_cleaned[, 2:25]) #NEED TO SKIP the first column of genes
class(exprsFull)
dim(exprsFull) #16002 genes X 24 samples (mice)
colnames(exprsFull)#MDST names
rownames(exprsFull)<-t(MDST_RNASeq_dataFrame_cleaned[,1])# not set? #added tues

head(exprsFull)

#create a minimal ExpressionSet object
mdstSet<-ExpressionSet(assayData=exprsFull[1:16002, 1:24]) #this works but not passing the full object??

#run gsva
gsva_mdst<-gsva(mdstSet,gset.idx.list=msigdbr_list, method=c("gsva"), verbose=TRUE) #testing 03/01/21
gsva_mdst

#extract th expression matrix -> should this now be 
mdst_MatrixES<-exprs(gsva_mdst)
View(mdst_MatrixES)

repStress_hallmarks<-as.data.frame(c("HALLMARK_G2M_CHECKPOINT", "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
                       "HALLMARK_E2F_TARGETS", "HALLMARK_DNA_REPAIR", 
                       "HALLMARK_MYC_TARGETS_V1", "HALLMARK_MYC_TARGETS_V2"))
colnames(repStress_hallmarks)<-c("Hallmark")
View(repStress_hallmarks)

#correlation of hallmark scores for controls
#label pathways of interest
#controls_ES<-as.data.frame(mdst_MatrixES)%>%
#  dplyr::select(Control1_FatPad_F1, Control1_FatPad_F2)
#View(controls_ES)

forScatterPlot<-as.data.frame(mdst_MatrixES) %>%
  rownames_to_column("Hallmark") %>%
  mutate(RepStress_Hallmark=ifelse((Hallmark %in% repStress_hallmarks$Hallmark), "Yes", "No"))

View(forScatterPlot)

scatter_GSVA_controls<-forScatterPlot%>%
  ggscatter(x = "Control1_FatPad_F1", y = "Control1_FatPad_F2", color="RepStress_Hallmark",
            #label="Hallmark",repel=TRUE,
            add = "reg.line", 
            add.params = list(color = "dodgerblue2", fill = "lightgray"),
            conf.int = FALSE, title = "Correlation of GSVA enrichment scores", subtitle = "for MDST Models- Controls",
            xlab = "Control1 FatPad F1 ", ylab = "Control1 FatPad F2", 
            #label.select = ifelse(RepStress_Hallmark=="Yes"),
            font.label = c(6, "bold")) + 
  stat_cor(method = "spearman",color = "dodgerblue2", size = 4.5, label.x=0.30, label.y = 0.75)
scatter_GSVA_controls

#facet(scatter_CyclinE, facet.by = "Signature")

```

```{r}
#SET UP FOR HEATMAP

#just our subset of hallmarks here
forHeatmap_RepStress_HallMarks<-as.data.frame(mdst_MatrixES) %>%
  rownames_to_column("Hallmark") %>%
  mutate(RepStress_Hallmark=ifelse((Hallmark %in% repStress_hallmarks$Hallmark), "Yes", "No"))%>%
  filter(RepStress_Hallmark=="Yes")#%>%
  #mutate(Hallmark_Cl = str_sub(Hallmark, 10, length(Hallmark))) #not working?
View(forHeatmap_RepStress_HallMarks)

#set up for heatmap
removeHallmarkNames<-forHeatmap_RepStress_HallMarks %>%
  dplyr::select(-Hallmark, -RepStress_Hallmark)
#View(removeHallmarkNames)
matrix_forHeatMap<-as.matrix(removeHallmarkNames)
rownames(matrix_forHeatMap)<-forHeatmap_RepStress_HallMarks$Hallmark
```

```{r}
#heatmap 
png(filename = here("output", "GSVA", "MDST_GSVA_RepStressHallmarks.png"), width = 1200, height = 400)
#works but branches small
htmp_rsr<-Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Enrichment Score", 
        #right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 8, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        column_dend_height = unit(2, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
draw(htmp_rsr, heatmap_legend_side = "left", annotation_legend_side = "bottom")
dev.off()
```
```{r}
#HEATMAP GSVA MDST -  ALL

#just our subset of hallmarks here
forHeatmap_RepStress_HallMarksAll<-as.data.frame(mdst_MatrixES) %>%
  rownames_to_column("Hallmark")%>%
  mutate(Hallmark = str_sub(Hallmark, 10, length(Hallmark)))

View(forHeatmap_RepStress_HallMarksAll)

#set up for heatmap
removeHallmarkNames<-forHeatmap_RepStress_HallMarksAll %>%
  dplyr::select(-Hallmark)
#View(removeHallmarkNames)
matrix_forHeatMap<-as.matrix(removeHallmarkNames)
rownames(matrix_forHeatMap)<-forHeatmap_RepStress_HallMarksAll$Hallmark
```

```{r}
#heatmap 
png(filename = here("output", "GSVA", "MDST_GSVA_AllHallmarks.png"), width = 1800, height =1200)
#works but branches small
htmp<-Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Enrichment Score", 
        #right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 8.5, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        column_dend_height = unit(3, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
draw(htmp, heatmap_legend_side = "left", annotation_legend_side = "bottom")
dev.off()
```

```{r}
#COMPARE GSVA HALLMARKS REP STRESS WITH RSR SCORE (and PARPness score)

#join with RSR mcgrail scores
#these are z-scores, and cyclin E log2cpm* should we z-score this
MDST_Zscores<-read_tsv(here("data_intermediate", "MDST", "MDST_RSR_PARPness_CCE1_WIDE.tsv"))
View(MDST_Zscores)

#long format score (hallmark)
#correlation is hallmark vs RSR
#correlation of each hallmark and RSR score
#facet by hallmark
#mutate(Hallmark_Cl = str_sub(Hallmark, 10, length(Hallmark))) #not working?
View(forHeatmap_RepStress_HallMarks)
#need to re-arrange here
ModelsToHallmarks<-forHeatmap_RepStress_HallMarks %>%
  select(-RepStress_Hallmark) %>%
  pivot_longer(-Hallmark, names_to = "Model", values_to="Value") %>%
  pivot_wider(names_from = Hallmark, values_from = Value)
View(ModelsToHallmarks)

joinedMDST_Scores<-left_join(ModelsToHallmarks, MDST_Zscores)
View(joinedMDST_Scores)

#format just for GSVA versus RSR/PARPness scores
#get hallmarks-> wide | keep model and mcgrail RSR/PARPness/CCNE1 info in own columns (will be doubled up)
#then make hallmarks WIDE so we can iterate through/facet in our plots
joinedMDST_Scores_forScatter<-joinedMDST_Scores %>%
  pivot_longer(-c(Model, McGrail_RSR, McGrail_PARPness, Ccne1_Log2CPM) , names_to= "Hallmark", values_to = "ES")
View(joinedMDST_Scores_forScatter)

#MDST RSR versus GSVA sets of interest
scatter_GSVA_vsRSR<-joinedMDST_Scores_forScatter%>%
  ggscatter(x = "McGrail_RSR", y = "ES", #color="RepStress_Hallmark",
            #label="Hallmark",repel=TRUE,
            add = "reg.line", 
            add.params = list(color = "dodgerblue2", fill = "lightgray"),
            conf.int = FALSE, title = "Correlation of GSVA Scores with  McGrail RSR", subtitle = "MDST Models",
            xlab = "McGrail RSR Z-Score", ylab = "HallMark ES", 
            font.label = c(6, "bold")) + 
  stat_cor(method = "spearman",color = "dodgerblue2", size = 2.5, label.x=0.0, label.y = 0.75)
#scatter_GSVA_vsRSR
facet(scatter_GSVA_vsRSR, facet.by = "Hallmark")

#MDST PARPness versus GSVA sets of interest
scatter_GSVA_vsPARPness<-joinedMDST_Scores_forScatter%>%
  ggscatter(x = "McGrail_PARPness", y = "ES", #color="RepStress_Hallmark",
            #label="Hallmark",repel=TRUE,
            add = "reg.line", 
            add.params = list(color = "dodgerblue2", fill = "lightgray"),
            conf.int = FALSE, title = "Correlation of GSVA Scores with  McGrail PARPness", subtitle = "MDST Models",
            xlab = "McGrail PARPness Z-Score", ylab = "HallMark ES", 
            font.label = c(6, "bold")) + 
  stat_cor(method = "spearman",color = "dodgerblue2", size = 2.5, label.x=0.0, label.y = 0.75)
#scatter_GSVA_vsPARPness
facet(scatter_GSVA_vsPARPness, facet.by = "Hallmark")

#MDST cyclin E vs GSVA sets of interest
scatter_GSVA_vsCCNE1<-joinedMDST_Scores_forScatter%>%
  ggscatter(x = "Ccne1_Log2CPM", y = "ES", #color="RepStress_Hallmark",
            #label="Hallmark",repel=TRUE,
            add = "reg.line", 
            add.params = list(color = "dodgerblue2", fill = "lightgray"),
            conf.int = FALSE, title = "Correlation of GSVA Scores with  Cyclin E Expression", subtitle = "MDST Models",
            xlab = "CCCNE1 Log2(CPM)", ylab = "HallMark ES", 
            font.label = c(6, "bold")) + 
  stat_cor(method = "spearman",color = "dodgerblue2", size = 2.5, label.x=0.0, label.y = 0.75)
#scatter_GSVA_vsCCNE1
facet(scatter_GSVA_vsCCNE1, facet.by = "Hallmark")
```

```{r}
#Updated heatmap GSVA RSR - add scores and cyclin E
#View(joinedMDST_Scores) #use this data frame, hallmarks plus scores


######
#set up for heatmap START HERE*
removeNames<-joinedMDST_Scores %>%
  dplyr::select(-Model, -McGrail_RSR, -McGrail_PARPness, -Ccne1_Log2CPM)
#View(removeHallmarkNames)
matrix_forHeatMap<-as.matrix(removeNames)
rownames(matrix_forHeatMap)<-joinedMDST_Scores$Model

col_RSR = colorRamp2(c(-4, 0,4), c("turquoise1","wheat","darkorange1"))
col_PARP = colorRamp2(c(-4,0,4), c("turquoise1","wheat","darkorange1"))
col_CCNE1 = colorRamp2(c(0,6), c("forestgreen","orange1"))

#row annotation
#McGrail RSR | McGrail PARPness  | CyclinE
row_ha = rowAnnotation(Mc_RSR = joinedMDST_Scores$McGrail_RSR, 
                       Mc_PARP = joinedMDST_Scores$McGrail_PARPness,
                       CCNE1 = joinedMDST_Scores$Ccne1_Log2CPM,
                       col=list(Mc_RSR = col_RSR, 
                                Mc_PARP = col_PARP,
                                CCNE1=col_CCNE1)
                       )

```

```{r}
#heatmap 
png(filename = here("output", "GSVA", "MDST_GSVA_RSRHallmarks_AddScores.png"), width = 1000, height =1400)
#works but branches small
htmp_withScores<-Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Enrichment Score", 
        right_annotation = row_ha, 
        row_names_gp = gpar(fontsize = 12, fontface = "bold"),
        column_names_gp= gpar(fontsize =12, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        column_dend_height = unit(2, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
draw(htmp_withScores, heatmap_legend_side = "left", annotation_legend_side = "bottom")
dev.off()
```

```{r}
#FIX ORIENTATION OF HEATMAP
#quick try - fix orientation to match gsva heatmap
View(joinedMDST_Scores)
changeOrientation<-joinedMDST_Scores %>%
  pivot_longer(-Model) %>%
  pivot_wider(names_from= Model, values_from = value)
View(changeOrientation)

matrixOnly<-changeOrientation[1:6,]
View(matrixOnly)

#set up for heatmap
removeHallmarkNames<-matrixOnly%>%
  dplyr::select(-name)
#View(removeHallmarkNames)
matrix_forHeatMap<-as.matrix(removeHallmarkNames)
rownames(matrix_forHeatMap)<-matrixOnly$name

#problem - we need rows to be our annotation, may have to grab each one* kind of a pain
Mc_RSR <-as.data.frame(changeOrientation[8,2:25])
View(t(Mc_RSR))

Mc_PARP <-as.data.frame(changeOrientation[9,2:25])
View(t(Mc_PARP))

CCNE1 <-as.data.frame(changeOrientation[7,2:25])
View(t(CCNE1))

#color palettes
col_RSR = colorRamp2(c(-4, 0,4), c("turquoise4","grey75","darkorange1"))
col_PARP = colorRamp2(c(-4,0,4), c("turquoise4","grey75","darkorange1"))
col_CCNE1 = colorRamp2(c(0,6), c("forestgreen","orange1"))

#
#column annotation in this orientation
col_ha = HeatmapAnnotation(McGrail_RSR = t(Mc_RSR), 
                         McGrail_PARP = t(Mc_PARP),
                         CCNE1_Exp = t(CCNE1),
                          col=list(McGrail_RSR  = col_RSR, 
                                McGrail_PARP = col_PARP,
                                CCNE1_Exp=col_CCNE1)
                       )
```

```{r}
#heatmap 
png(filename = here("output", "GSVA", "MDST_GSVA_RepStressHallmarks_AddScores_Orientation.png"), width = 1400, height = 500)
#works but branches small
htmp_rsr<-Heatmap(matrix_forHeatMap, 
        cluster_rows = TRUE,
        cluster_columns = TRUE, 
        show_column_names = TRUE, 
        show_row_names = TRUE,
        name= "Enrichment Score", 
        bottom_annotation = col_ha, 
        row_names_gp = gpar(fontsize = 8, fontface = "bold"),
        column_names_gp= gpar(fontsize =16, fontface = "bold"),
        row_dend_width = unit(4, "cm"),
        column_dend_height = unit(2, "cm")
        #width = unit(72, "cm"), #heatmap body width
        #height = unit(56, "cm") ) #heatmap body height
        )
draw(htmp_rsr, heatmap_legend_side = "left", annotation_legend_side = "left")
dev.off()
```




```