---
title: "RepStress Project - Select Cell Lines"
author: "Aurora S Blucher"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(caret)
```

## Split DepMap/CCLE Lines -> Training/Test Sets

This script reads in the DepMap/CCLE data that we have and splits into an 80/20 training and test set. Write sets to different files. 
For breast, ovarian, endometrial, and pancreatic lines (n=244 total). Split and keep an equal proportion of each disease type. 

Notes
02/02/21 - read in and split

Citations
-tidyverse; Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686
-here
-readxl
-caret; https://topepo.github.io/caret/data-splitting.html

```{r}
#Cell lines 01/28/20 file from DDR-Pathway-Resources>output>depmap>queriese
Lines_FullSet<-read_tsv(here("data", "DepMap_forRepStress_BrEndoOvPancr_012821.txt"))
View(Lines_FullSet) #244 lines total

#random seed
set.seed(020221)

#split data into 80/20; with respect to CancerType; list = FALSE = don't return data as a ilist
#use createResampple to do bootstraps; createFolds to to do balanced cross-validation groupings
#cite: Max Kuhn/ caret package https://topepo.github.io/caret/data-splitting.html
trainIndex<-createDataPartition(Lines_FullSet$Cancer_Type, p=.8, list = FALSE, times = 1)
head(trainIndex)

Lines_Training<-Lines_FullSet[trainIndex,] #take rows from our index, all columns
Lines_Test<-Lines_FullSet[-trainIndex, ] #take all other rows, all columns

#quick check #s
#TRAINING = 197
View(Lines_Training)
write_tsv(Lines_Training, here("data", "CCLE_Training", "DepMap_forRepStress_BrEndoOvPanc_TRAINING_020221.tsv"))
#TEST = 47
View(Lines_Test)
write_tsv(Lines_Test, here("data", "CCLE_Test", "DepMap_forRepStress_BrEndoOvPanc_TEST_020221.tsv"))


```