---
title: "SplitCCLE_Training_Test_AllOtherDiseases"
author: "Aurora S Blucher"
date: "4/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(here)
library(caret)
```

## Splitting Full CCLE (All Diseases) -> Training and Test

This is an additional script for splitting the full CCLE data into training and test sets. First we need to pull off breast/ovarian/endometrial/pancreatic lines, which have previously been split; then select a random 80% of the lines for training.  Require an equal distribution by disease type. 

Notes
04/20/20 - split other diseases into training/ test, run the McGrail RSR RNA signature


```{r}
#read in the full list of cell lines

CCLE_all<-read_csv(here("data", "DepMapCCLE_data", "sample_info_cleaned_021521.csv"))
View(CCLE_all)

ExcludeDiseases<-as.data.frame(c("Breast Cancer", "Ovarian Cancer", "Endometrial/Uterine Cancer", "Pancreatic Cancer"))
colnames(ExcludeDiseases)<-c("Disease")
#remove any without disease or more information
#remove breast, ovarian, endo, pancreatic
CCLE_additionalDiseases<-CCLE_all %>%
  filter(!primary_disease %in% ExcludeDiseases$Disease) %>%
  filter( !is.na(cell_line_name)) %>%
  filter(primary_disease!="Unknown")
View(CCLE_additionalDiseases)

#random seed
set.seed(042021)

#SPLIT INTO TRAINING/TEST SETS
#split data into 80/20; with respect to CancerType; list = FALSE = don't return data as a ilist
#use createResampple to do bootstraps; createFolds to to do balanced cross-validation groupings
#cite: Max Kuhn/ caret package https://topepo.github.io/caret/data-splitting.html
trainIndex<-createDataPartition(CCLE_additionalDiseases$primary_disease, p=.8, list = FALSE, times = 1)
head(trainIndex)

#subset
Lines_Training<-CCLE_additionalDiseases[trainIndex,] #take rows from our index, all columns
Lines_Test<-CCLE_additionalDiseases[-trainIndex, ] #take all other rows, all columns

#quick check #s
#TRAINING =1150
View(Lines_Training)
#write_tsv(Lines_Training, here("data", "CCLE_Training", "DepMap_forRepStress_OtherDiseasess_TRAINING_210421.tsv"))
#TEST = 274
View(Lines_Test)
#write_tsv(Lines_Test, here("data", "CCLE_Test", "DepMap_forRepStress_OtherDiseases_TEST_210421.tsv"))

```

```{r}
#TRAINING only
CCLE_allOtherDiseases<-read_tsv(here("data", "CCLE_Training", "DepMap_forRepStress_OtherDiseasess_TRAINING_210421.tsv")) %>%
  select(DepMapID=DepMap_ID, CellLine_Name = cell_line_name, CellLine_Name_Cleaned = stripped_cell_line_name, 
         Cancer_Type = primary_disease, Cancer_Subtype = Subtype, DerivedFrom_Primary_or_Met=primary_or_metastasis )
View(CCLE_allOtherDiseases)
#TRAINING - breast, ov, endo, pancreatic from main scripting work
#merge
CCLE_BrOvEndPanc<-read_tsv(here("data", "CCLE_Training", "DepMap_forRepStress_BrEndoOvPanc_TRAINING_020221.tsv"))
View(CCLE_BrOvEndPanc)

mergedFrames<-rbind(CCLE_allOtherDiseases, CCLE_BrOvEndPanc)
View(mergedFrames)

#RNASeq data
#go back to full CCLE file
DepMap_RNASeq<-read_csv(here("data", "DepMapCCLE_data", "CCLE_expression.csv"))
View(DepMap_RNASeq)

#Filter RNASeq data down to our cell lines; 
#cell lines
DepMap_TrainingLines_RNASeq_All<-DepMap_RNASeq %>%
  pivot_longer(-X1, names_to = "Gene", values_to = "Log2_TPM") %>%
  select(DepMap_ID = X1, Gene, Log2_TPM) %>%
  filter(DepMap_ID %in% mergedFrames$DepMapID) %>% #need to use DepMapID across data types
  separate(Gene, c("Name", "ID_ToClean"), " ") %>%
  mutate(ID = str_sub(ID_ToClean, 2, -2)) %>%#start at second char, end on 2nd to last chat -> remove "()"
  select(DepMap_ID, Gene= Name, Gene_ID = ID, Log2_TPM)
dim(DepMap_TrainingLines_RNASeq_All) #20371284 x 4
#View(DepMap_TrainingLines_RNASeq)
#intermediate, output to file
write_tsv(DepMap_TrainingLines_RNASeq_All, here("data_intermediate", "CCLEDepMap_expression", "CCLEDepMap_TrainingLines_AllDiseases_RNASeq_LONG.tsv"))


#McGrail RSR Genes and Coefficients
RSRSignatureGenes<-read_tsv(here("data","RNA_Signatures", "McGrailRSR_Genes_Coefficients.tsv")) %>%
  arrange(Gene)
View(RSRSignatureGenes)

#now need to clean the Gene field and filter to our genes
DepMap_RNASeq_RSRGenes<-DepMap_TrainingLines_RNASeq_All %>%
 filter( Gene %in% RSRSignatureGenes$Gene) 
dim(DepMap_RNASeq_RSRGenes)
View(unique(DepMap_RNASeq_RSRGenes$Gene)) #580 genes
#View(DepMap_RNASeq_RSRGenes)

RSRSignatureGenes_filtered<-RSRSignatureGenes %>%
   filter(Gene %in% DepMap_RNASeq_RSRGenes$Gene)
#View(RSRSignatureGenes_filtered)

#re-structure, then run against corelation of RSR
DepMap_RNASeq_RSRGenes_cleaned_dataframe<-DepMap_RNASeq_RSRGenes %>%
  select(DepMapID = DepMap_ID, Gene, Log2_TPM ) %>%
  arrange(Gene) 
#View(DepMap_RNASeq_RSRGenes_cleaned_dataframe)

#join with our gene co-efficients  
Joined_DepMapRNASeq_RSRCoefficients<-left_join(DepMap_RNASeq_RSRGenes_cleaned_dataframe,RSRSignatureGenes_filtered )
#View(Joined_DepMapRNASeq_RSRCoefficients)

#run correlations on full data frame
dataFrame_corr_all<-Joined_DepMapRNASeq_RSRCoefficients %>%
  group_by(DepMapID) %>%
  summarize(correlation = cor(Log2_TPM, Coefficient )) 
View(dataFrame_corr_all)
#intermediate output correlations; these can go in the intermediate files
#write_tsv(dataFrame_corr_all, here("data_intermediate", "McGrail_RSR", "CCLEDepMap_TrainingLines_AllDiseases_RSR.tsv"))

#join back with cell line info
joinRSRCorr_CellLineInfo<-left_join(dataFrame_corr_all, mergedFrames) %>%
  select(DepMapID, correlation, CellLine_Name_Cleaned, Cancer_Type, Cancer_Subtype, DerivedFrom_Primary_or_Met)
View(joinRSRCorr_CellLineInfo)
#write_tsv(joinRSRCorr_CellLineInfo, here("data_intermediate", "McGrail_RSR", "CCLEDepMap_TrainingLines_AllDiseases_RSR_JoinedCellLineInfo.tsv"))

```

```{r}
#Split into separate script with clean up*
#Read in data and run the McGrail RSR
readInData<-read_tsv(here("data_intermediate", "McGrail_RSR", "CCLEDepMap_TrainingLines_AllDiseases_RSR_JoinedCellLineInfo.tsv"))
View(readInData)

#z-score transform
BindScores<-readInData %>%
  mutate(Score = "McGrail_RSR") %>%
  group_by(Score) %>%
  filter(!is.na(correlation)) %>% #make sure to remove the NAs
  mutate(Mean = mean(correlation), SD = sd(correlation), ZScore= (correlation- Mean)/SD) %>%
  dplyr::select(-Mean, -SD, -correlation) %>%
  pivot_wider(names_from = Score, values_from = ZScore) #checked, looks good
View(BindScores)

#graph all diseases RSR McGrail - distribution here
plotSubtype<- BindScores%>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR)) +
  geom_bar(stat="identity", aes(fill=Cancer_Type)) #+ 
  #scale_fill_manual(values = cols, na.value = "grey")  
plotFinal<-plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE All - Training (80%)") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(#axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x=element_blank(),
        axis.ticks.x =element_blank(),
        axis.text.x=element_blank()) 
plotFinal
ggsave(here("output", "TestHeatmap_All.png"), width = 26, height = 20, units = c("cm"), dpi = 300)

#zoom in to just the high scoring RSR scores
#zscore over 2
plotSubtype<- BindScores%>%
  filter(McGrail_RSR>=2.0) %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR)) +
  geom_bar(stat="identity", aes(fill=Cancer_Type)) #+ 
  #scale_fill_manual(values = cols, na.value = "grey")  
plotFinal<-plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE All - Training (80%)") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(#axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x=element_blank(),
        axis.ticks.x =element_blank(),
        axis.text.x=element_blank()) 
plotFinal
ggsave(here("output", "TestHeatmap_All_HighRSR.png"), width = 26, height = 20, units = c("cm"), dpi = 300)

#z-score over 1.5
plotSubtype<- BindScores%>%
  filter(McGrail_RSR>=1.5) %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR)) +
  geom_bar(stat="identity", aes(fill=Cancer_Type)) #+ 
  #scale_fill_manual(values = cols, na.value = "grey")  
plotFinal<-plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE All - Training (80%)") +
  xlab("") +
  ylab("Z-Score") + 
  guides(fill=guide_legend(title="Subtype")) + 
  theme(#axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x=element_blank(),
        axis.ticks.x =element_blank(),
        axis.text.x=element_blank()) +
  facet_wrap(~Cancer_Type)
plotFinal
ggsave(here("output", "TestHeatmap_All_HighRSR_1Pt5_Facet.png"), width = 26, height = 20, units = c("cm"), dpi = 300)



```

```{r}
#EXAMINE A FEW DISEASES OF INTERESET
plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Breast Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Breast") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#ovarian
#EXAMINE A FEW DISEASES OF INTERESET
plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Ovarian Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Ovarian") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

#endo
#EXAMINE A FEW DISEASES OF INTERESET
plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Endometrial/Uterine Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Endometrial") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Pancreatic Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Pancreatic") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Head and Neck Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Head and Neck") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


plotSubtype<- BindScores%>%
  filter(Cancer_Type== "Bladder Cancer") %>%
  arrange(desc(McGrail_RSR)) %>%
  ggplot(aes(x=reorder(CellLine_Name_Cleaned, -McGrail_RSR), y=McGrail_RSR),
         color="black") +
  geom_bar(stat="identity", aes(fill=McGrail_RSR)) + #, color=BRCA_Status)) + 
  scale_fill_continuous(low = "dodgerblue2", high = "tomato")
plotSubtype + labs(title = "Distribution of McGrail RSR Score (Z-Scores)", subtitle = "CCLE Bladder") +
  xlab("") +
  ylab("Z-Score") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


